---
title: "R Notebook"
output: html_notebook
---

```{r echo=FALSE, message=FALSE}
require('tidyverse')
require('ggridges')
require('GGally')
require('cowplot')

```



# Import Data
```{r}
wn_train <- read.csv('data/train.csv', header=TRUE, stringsAsFactors = FALSE)
wn_test <- read.csv('data/test.csv', header=TRUE, stringsAsFactors = FALSE)

weather <- read.csv('data/weather.csv', header=TRUE,na.strings = 'M', stringsAsFactors = FALSE)
spray <- read.csv('data/spray.csv', header=TRUE, stringsAsFactors = FALSE)
```




# Inital Clean-up
```{r, message=FALSE, warning=FALSE}

spray <- unique(spray[,c('Date', 'Latitude', 'Longitude')]) %>%
       mutate(
              dayid = as.integer(gsub('[^0-9]','', Date))
              , year = as.integer(substr(dayid, 1,4))
              , month = as.integer(substr(dayid, 5,6))
              , day = as.integer(substr(dayid, 7,8))
              , Latitude = round(Latitude, 3) # roughly nearest 300ft
              , Longitude = round(Longitude, 3) # roughly nearest 300ft
       ) %>%
       group_by(year, month, Latitude, Longitude) %>%
       count() %>%
       select(-n) %>%
       rename(spray_lat = Latitude
              , spray_long = Longitude)

df_train <- wn_train %>% 
       group_by(Date, Species, Trap, Latitude, Longitude) %>%
       summarize(
              NumMosquitos = sum(NumMosquitos)
              , WnvPresent = max(WnvPresent)
       ) %>%
       mutate(
              dayid = as.integer(gsub('[^0-9]','', Date))
              , year = as.integer(substr(dayid, 1,4))
              , month = as.integer(substr(dayid, 5,6))
              , day = as.integer(substr(dayid, 7,8))
              , isTrain = 1
       )  %>%
       ungroup() %>%
       rownames_to_column(var='Id')

df_test <- wn_test %>% 
       mutate(
              Trap_adjusted = ifelse(grepl('[^0-9]$', Trap, ignore.case=TRUE), sub('A$','',Trap),Trap)
              , dayid = as.integer(gsub('[^0-9]','', Date))
              , year = as.integer(substr(dayid, 1,4))
              , month = as.integer(substr(dayid, 5,6))
              , day = as.integer(substr(dayid, 7,8))
              , isTrain = 0
              , WnvPresent = 0
       )

year_monthly_weather <- weather %>%
       mutate(
            dayid = as.integer(gsub('[^0-9]','', Date))  
            , year = as.integer(substr(dayid, 1,4))
            , month = as.integer(substr(dayid, 5,6))
       ) %>% 
       group_by(year, month) %>%
       summarize(
              avg_temp = mean(Tavg, na.rm=TRUE)
              , max_temp = max(Tmax, na.rm=TRUE)
              , min_temp = min(Tmin, na.rm=TRUE)
              #, mean_depart = mean(Depart, na.rm=TRUE)
              , PrecipTotal = sum(as.numeric(ifelse(PrecipTotal=='T',0,PrecipTotal)), na.rm=TRUE)
              , wind_speed = mean(AvgSpeed, na.rm=TRUE)
       )

monthly_weather <- weather %>%
       mutate(
            dayid = as.integer(gsub('[^0-9]','', Date))  
            , year = as.integer(substr(dayid, 1,4))
            , month = as.integer(substr(dayid, 5,6))
       ) %>% 
       group_by(month) %>%
       summarize(
              avg_temp = mean(Tavg, na.rm=TRUE)
              , max_temp = max(Tmax, na.rm=TRUE)
              , min_temp = min(Tmin, na.rm=TRUE)
              #, mean_depart = mean(Depart, na.rm=TRUE)
              , PrecipTotal = sum(as.numeric(ifelse(PrecipTotal=='T',0,PrecipTotal)), na.rm=TRUE)
              , wind_speed = mean(AvgSpeed, na.rm=TRUE)
       )

# daily_weather <- weather %>%
#        mutate(
#             dayid = as.integer(gsub('[^0-9]','', Date))  
#             , year = as.integer(substr(dayid, 1,4))
#             , month = as.integer(substr(dayid, 5,6))
#        ) %>% 
#        group_by(dayid) %>%
#        summarize(
#               avg_temp = mean(Tavg, na.rm=TRUE)
#               , max_temp = max(Tmax, na.rm=TRUE)
#               , min_temp = min(Tmin, na.rm=TRUE)
#               , mean_depart = mean(Depart, na.rm=TRUE)
#               , PrecipTotal = sum(as.numeric(ifelse(PrecipTotal=='T',0,PrecipTotal)), na.rm=TRUE)
#               , wind_speed = mean(AvgSpeed, na.rm=TRUE)
#        )

```




# Initial EDA Plots

## Exploring Year / Month / Year + Month impact

### Year
- Deploying less traps per year, but more efficient in placement
- 2009 / 2011 are interesting, lower rates.

```{r}

yearly_summary <- df_train %>%
       group_by(year) %>%
       summarize(
              num_samples = n()
              , distinct_Traps = n_distinct(Trap)
              , distinct_Wn_Traps = n_distinct(ifelse(WnvPresent>0,Trap,1)) - 1
              , percent_positive_traps = distinct_Wn_Traps / distinct_Traps
              , samples_per_Trap = n() /  n_distinct(Trap)
              , west_nile_rate = mean(WnvPresent)
              , avg_mosquitoes = mean(NumMosquitos)
              , total_mosquitoes = sum(NumMosquitos)
       ) %>% 
       arrange(year)

yearly_summary
```

### Monthly
- Highest rates in August.  
```{r}

monthly_summary <- df_train %>%
       group_by(month) %>%
       summarize(
              num_samples = n()
              , distinct_Traps = n_distinct(Trap)
              , distinct_Wn_Traps = n_distinct(ifelse(WnvPresent>0,Trap,1)) - 1
              , percent_positive_traps = distinct_Wn_Traps / distinct_Traps
              , samples_per_Trap = n() /  n_distinct(Trap)
              , west_nile_rate = mean(WnvPresent)
              , avg_mosquitoes = mean(NumMosquitos)
              , total_mosquitoes = sum(NumMosquitos)
       ) %>% 
       arrange(month)

monthly_summary
```



```{r}
year_monthly_summary <- df_train %>%
       group_by(year,month) %>%
       summarize(
              num_samples = n()
              , distinct_Traps = n_distinct(Trap)
              , distinct_Wn_Traps = n_distinct(ifelse(WnvPresent>0,Trap,1)) - 1
              , percent_positive_traps = distinct_Wn_Traps / distinct_Traps
              , samples_per_Trap = n() /  n_distinct(Trap)
              , west_nile_rate = mean(WnvPresent)
              , avg_mosquitoes = mean(NumMosquitos)
              , total_mosquitoes = sum(NumMosquitos)
       ) %>% 
       arrange(year,month)

year_monthly_summary
```



```{r}


percent_positive_traps_plot <- ggplot(year_monthly_summary) +
       geom_tile(aes(x=year, y=month, fill=percent_positive_traps))

west_nile_rate_plot <- ggplot(year_monthly_summary) +
       geom_tile(aes(x=year, y=month, fill=west_nile_rate))

distinct_Traps_plot <- ggplot(year_monthly_summary) +
       geom_tile(aes(x=year, y=month, fill=distinct_Traps))

distinct_Wn_Traps_plot <- ggplot(year_monthly_summary) +
       geom_tile(aes(x=year, y=month, fill=distinct_Wn_Traps))


plot_grid(percent_positive_traps_plot
          , west_nile_rate_plot
          , distinct_Traps_plot
          , distinct_Wn_Traps_plot)

```


## Monthly Temperatures
```{r}

temp_increment <- 2
monthly_temp_summary <- df_train %>%
       left_join(daily_weather, by = "dayid") %>%
       mutate(
              avg_temp_bucket = floor(avg_temp /temp_increment)*temp_increment
       ) %>%
       group_by(month, avg_temp_bucket) %>%
       summarize(
              num_samples = n()
              , distinct_Traps = n_distinct(Trap)
              , distinct_Wn_Traps = n_distinct(ifelse(WnvPresent>0,Trap,1)) - 1
              , percent_positive_traps = distinct_Wn_Traps / distinct_Traps
              , samples_per_Trap = n() /  n_distinct(Trap)
              , west_nile_rate = mean(WnvPresent)
              , avg_mosquitoes = mean(NumMosquitos)
              , total_mosquitoes = sum(NumMosquitos)
       ) %>% 
       arrange(month, avg_temp_bucket)

monthly_temp_summary
```


```{r}
percent_positive_traps_plot_temp <- ggplot(monthly_temp_summary) +
       geom_tile(aes(x=month, y=avg_temp_bucket, fill=percent_positive_traps))

west_nile_rate_plot_temp <- ggplot(monthly_temp_summary) +
       geom_tile(aes(x=month, y=avg_temp_bucket, fill=west_nile_rate))

distinct_Traps_plot_temp <- ggplot(monthly_temp_summary) +
       geom_tile(aes(x=month, y=avg_temp_bucket, fill=distinct_Traps))

distinct_Wn_Traps_plot_temp <- ggplot(monthly_temp_summary) +
       geom_tile(aes(x=month, y=avg_temp_bucket, fill=distinct_Wn_Traps))


plot_grid(percent_positive_traps_plot_temp
          , west_nile_rate_plot_temp
          , distinct_Traps_plot_temp
          , distinct_Wn_Traps_plot_temp)
```


## Summary By Type of Mosquito

```{r}
species_summary <- df_train %>%
       group_by(Species) %>%
       summarise(
              num_samples = n()
              , distinct_Traps = n_distinct(Trap)
              , distinct_Wn_Traps = n_distinct(ifelse(WnvPresent>0,Trap,1)) - 1
              , percent_positive_traps = distinct_Wn_Traps / distinct_Traps
              , samples_per_Trap = n() /  n_distinct(Trap)
              , west_nile_rate = mean(WnvPresent)
              , avg_mosquitoes = mean(NumMosquitos)
              , total_mosquitoes = sum(NumMosquitos)
       )

species_summary

```

# Finding Nearest Traps

```{r}

nearest_variables <- c('Trap', 'Latitude', 'Longitude')

trap_locations <- unique(df_train[,nearest_variables]) %>% 
       union_all(unique(df_test[,nearest_variables])) %>%
       group_by(Trap) %>%
       summarise(
              Latitude = max(Latitude)
              , Longitude = max(Longitude)
       ) %>%
       mutate(dummy=1) %>%
       ungroup()


trap_distances <- trap_locations %>%
       left_join(trap_locations %>% 
                        rename(second_Trap = Trap, second_lat = 'Latitude', second_long = 'Longitude')
                 , by='dummy') %>%
       filter(Trap != second_Trap) %>%
       mutate(
              distance = sqrt((Latitude - second_lat)^2 + (Longitude - second_long)^2)
       ) %>%
       select(Trap, second_Trap, distance) %>%
       arrange(Trap, distance) %>%
       group_by(Trap) %>%
       mutate(
              rank = row_number()
       )
```



```{r}
distance_plot <- trap_distances %>%
       filter(rank <= 3) %>%
       mutate(
              distance_bucket = round(distance, 4)
              , top_n = 3
       ) %>% 
       union_all(
       trap_distances %>%
       filter(rank <= 5) %>%
       mutate(
              distance_bucket = round(distance, 4)
              , top_n = 5
       )
       ) %>% 
       union_all(
       trap_distances %>%
       filter(rank <= 10) %>%
       mutate(
              distance_bucket = round(distance, 4)
              , top_n = 10
       )
       ) %>% 
       union_all(
       trap_distances %>%
       filter(rank <= 15) %>%
       mutate(
              distance_bucket = round(distance, 4)
              , top_n = 15
       )
       )

ggplot(distance_plot) +
       geom_density(aes(x=distance_bucket, fill = as.factor(top_n)),alpha= 0.5)

```


```{r}
top_n_distance <- 5
quantile_to_use <- '90%'

top_n_distances <- trap_distances %>%
       filter(rank <= top_n_distance)
 
probs = c(0, 0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99, 1)
distance_quantiles <- quantile(unlist(top_n_distances[,'distance']), probs = probs)

nearest_neighbor_distance <- distance_quantiles[quantile_to_use][[1]]

trap_distances <- trap_distances %>%
       mutate(
              isNear = ifelse(distance <= nearest_neighbor_distance,1,0)
       )
 
paste0("Nearest Neighbors Radius: ", round(nearest_neighbor_distance*69, 2), " miles")

```






```{r}
colnames(df_train)
```


# Nearest Spray
Note: Only spray data for two years (both train)
Not sure this will be an easy feature to add
```{r}

train_spray <- df_train[,c("Species", "Trap", "dayid", "year", "month", "Latitude", "Longitude")] %>%
       inner_join(spray, by = c('year' = 'year', 'month'='month')) %>%
             mutate(
              distance = sqrt((Latitude - spray_lat)^2 + (Longitude - spray_long)^2)
       ) %>%
       filter(distance <= nearest_neighbor_distance) %>%
       group_by(Species, Trap, dayid) %>%
       summarize(
              num_near_spray = n()
              , nearest_spray = min(distance)
       )

test_spray <- df_test[,c("Species", "Trap", "dayid", "year", "month", "Latitude", "Longitude")] %>%
       inner_join(spray, by = c('year' = 'year', 'month'='month')) %>%
             mutate(
              distance = sqrt((Latitude - spray_lat)^2 + (Longitude - spray_long)^2)
       ) %>%
       filter(distance <= nearest_neighbor_distance) %>%
       group_by(Species, Trap, dayid) %>%
       summarize(
              num_near_spray = n()
              , nearest_spray = min(distance)
       )

```




# Process Features

Function to process raw data
```{r}
process_data <- function(df) {
       
       base_cols <- c("Id", "Date", "Species", "Trap", "dayid", "year", "month", "day", "isTrain", "WnvPresent")
       
       df_temp <- df[,base_cols] %>%
       left_join(trap_distances[trap_distances$isNear == 1,], by="Trap") %>%
       left_join(df_train[, c('Trap', 'dayid', 'year', 'month','day', 'NumMosquitos', 'WnvPresent')], by = c('second_Trap' = 'Trap')) %>%
       filter(dayid.x > dayid.y & abs(year.x - year.y) <= 2 & year.x != year.y)  %>%
       group_by(Id, Species, Trap, dayid.x, WnvPresent.x) %>%
       summarize(
              year = max(year.x, na.rm=TRUE)
              , month = max(month.x, na.rm=TRUE)
              , day = max(day.x, na.rm=TRUE)
              , num_near_readings = n()
              , num_near_Traps = n_distinct(second_Trap)
              , num_near_Wn_Traps = n_distinct(ifelse(WnvPresent.y>0, second_Trap,1)) - 1
              , pct_near_Wn_Traps = num_near_Wn_Traps / num_near_Traps
              , is_Wn_near = max(WnvPresent.y, na.rm=TRUE)
              , num_Wn_near_records = sum(WnvPresent.y, na.rm=TRUE)
              , pct_Wn_near = sum(WnvPresent.y, na.rm=TRUE) / n()
              , num_Wn_mosquitos = sum(ifelse(WnvPresent.y>0, NumMosquitos,0), na.rm=TRUE)
              , num_mosquitos = sum(NumMosquitos, na.rm=TRUE)
              , distances_nearest_Wn = min(ifelse(WnvPresent.y>0, distance,1), na.rm=TRUE)
              , Wn_previous_location = max(ifelse(second_Trap == Trap, WnvPresent.y,0), na.rm=TRUE)
       ) %>%
       ungroup() %>%
       rename(dayid = dayid.x, WnvPresent = WnvPresent.x) 
       
       mean_non_near_distance <- mean(trap_distances[trap_distances$isNear == 0,'distance'][[1]])
       
       df_temp_2 <- df[,base_cols] %>%
              left_join(df_temp[,c(setdiff(colnames(df_temp),base_cols), 'Id')], by='Id') %>%
              left_join(year_monthly_weather, by=c('year'='year', 'month'='month')) %>%
              replace_na(replace = list(
                  num_near_readings = 0
                  , num_near_Traps = 0
                  , num_near_Wn_Traps = 0
                  , pct_near_Wn_Traps = 0
                  , is_Wn_near = 0
                  , num_Wn_near_records = 0
                  , pct_Wn_near = 0
                  , num_Wn_mosquitos = 0
                  , num_mosquitos = 0
                  , distances_nearest_Wn = mean_non_near_distance
                  , Wn_previous_location = 0  
                  
              ))
       
       return(df_temp_2)
}
```





Process data and create feature data frames for modeling

```{r}
df_train_features <- process_data(df_train)

test_split_size <- floor(nrow(df_test) / 5)

first <- seq(1, test_split_size)
second <- seq(test_split_size +1, test_split_size*2)
third <- seq(test_split_size*2 +1, test_split_size*3)
fourth <- seq(test_split_size*3 +1, test_split_size*4)
fifth <- seq(test_split_size*4 +1, nrow(df_test))

df_test_features_1 <- process_data(df_test[first,])
df_test_features_2 <- process_data(df_test[second,])
df_test_features_3 <- process_data(df_test[third,])
df_test_features_4 <- process_data(df_test[fourth,])
df_test_features_5 <- process_data(df_test[fifth,])

df_test_features <- df_test_features_1 %>%
       union_all(df_test_features_2) %>%
       union_all(df_test_features_3) %>%
       union_all(df_test_features_4) %>%
       union_all(df_test_features_5)

rm(df_test_features_1)
rm(df_test_features_2)
rm(df_test_features_3)
rm(df_test_features_4)
rm(df_test_features_5)
       

```






# Modelling

Ideas:
1. Adaboost 


## Old Stuff
```{r}
base_cols <- c("Date", "Species", "Trap", "dayid", "year", "month", "day", "isTrain")

df_train_features <- df_train[,base_cols] %>%
       #union_all(df_test[,base_cols]) %>%
       left_join(trap_distances[trap_distances$isNear == 1,], by="Trap") %>%
       left_join(df_train[, c('Trap', 'dayid', 'year', 'month','day', 'NumMosquitos', 'WnvPresent')], by = c('second_Trap' = 'Trap')) %>%
       filter(dayid.x > dayid.y, abs(year.x - year.y) <=2, year.x != year.y) %>%
       group_by(Species, Trap, dayid.x) %>%
       summarize(
              year = max(year.x)
              , month = max(month.x)
              , day = max(day.x)
              , num_near_readings = n()
              , num_near_Traps = n_distinct(second_Trap)
              , num_near_Wn_Traps = n_distinct(ifelse(WnvPresent>0, second_Trap,1)) - 1
              , pct_near_Wn_Traps = num_near_Wn_Traps / num_near_Traps
              , is_Wn_near = max(WnvPresent)
              , num_Wn_near_records = sum(WnvPresent)
              , pct_Wn_near = sum(WnvPresent) / n()
              , num_Wn_mosquitos = sum(ifelse(WnvPresent>0, NumMosquitos,0))
              , num_mosquitos = sum(NumMosquitos)
              , distances_nearest_Wn = min(ifelse(WnvPresent>0, distance,1))
              , Wn_previous_location = max(ifelse(second_Trap == Trap, WnvPresent,0))
       ) %>%
       ungroup() %>%
       rename(dayid = dayid.x) %>%
       left_join(year_monthly_weather, by=c('year'='year', 'month'='month'))




#year_monthly_weather

```



## Finding the most recent nearest West Nile trap

```{r}
nearest_variables <- c('Trap','dayid', 'year', 'Latitude', 'Longitude')
west_nile_location <- unique(df_train[df_train$WnvPresent == 1,c(nearest_variables, 'NumMosquitos')])
colnames(west_nile_location) <- paste0('wn_', c(nearest_variables, 'NumMosquitos'))
 
west_nile_distances <- unique(df_train[,nearest_variables]) %>% 
       union_all(unique(df_test[,nearest_variables])) %>%
       mutate(dummy = 1) %>%
       left_join(west_nile_location %>% mutate(dummy = 1), by = 'dummy') %>%
       filter(wn_dayid <= dayid, Trap != wn_Trap) %>%
       group_by(Trap,dayid,year, wn_Trap) %>%
       summarize(
              max_wn_dayid = max(wn_dayid) 
              , max_wn_year = max(wn_year) 
              , Latitude = max(Latitude)
              , Longitude = max(Longitude)
              , wn_Latitude = max(wn_Latitude)
              , wn_Longitude = max(wn_Longitude)
       ) %>%
        mutate(
               distance = sqrt((Latitude - wn_Latitude)^2 + (Longitude - wn_Longitude)^2)
        ) %>%
       left_join(west_nile_location[,c('wn_dayid', 'wn_Trap', 'wn_NumMosquitos')]
              , by = c('wn_Trap' = 'wn_Trap', 'max_wn_dayid' = 'wn_dayid')) 
       



```


## Stats for nearest Trap with West Nile that occured BEFORE the current test
```{r}
nearest_west_nile <- west_nile_distances %>%
       group_by(Trap, dayid) %>%
       top_n(1,desc(distance))
```

# Find most recent Spray time per location
```{r}

```




